---
title: "Initial Korean covid 19"
author: "Aetan Barquero"
date: "Fall 2025"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?






## Graphical results - use ggplot commands to do this  

```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=result,y=count,fill=Sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=result,y=prcnt*100,fill=Sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

  
By looking at the graphical plot, I noticed that deceased men are exceedingly higher than deceased women. The number which men die to covid compared to men is about a 2:1 ratio. 2/3 of the dead people were men. I also observed that the number of people alive is much larger than the amount of people dead. Most of the alive people were woman (about 57%) compared to men (43%). 




  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

The numerical results display the percentage of people that caught covid and either lived or dies. 1.4% of females that caught covid passed away and 3% of males that caught covid passed away.

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

After observing the inferential results, I concluded that the odds of a man to die to covid in South Korea are 2.3% greater than the odds of a female dying to covid in South Korea. This supports the previous results because we saw that the mortality rate to covid was higher in men. I also noticed that the p_value is extremely low at a value of 0.0009 thus meaning the chances of dying to COVID as a male is higher than a female's chance.

## Conclusions

In this assignment I analyzed the data of COVID in South Korea. Since I recognized that the act of dying or living and being male or female are both categorical, I determined to use the categorical-categoricla helper file to guide me through this. In the graphical results I was bale to do a chart with gg plot and observed that deceased men are exceedingly higher than deceased women. The number which men die to covid compared to men is about a 2:1 ratio. 2/3 of the dead people were men. I also observed that the number of people alive is much larger than the amount of people dead. Most of the alive people were woman (about 57%) compared to men (43%). In the numerical category, the results supported the previous results due to recording percentage and not count. The numerical results display the percentage of people that caught covid and either lived or dies. 1.4% of females that caught covid passed away and 3% of males that caught covid passed away. And the inferential results helped me calculate the odds and difference in odds of a male dying to covid compared to a female dying to covid.  
 
